---
title: "01_data_quality_checks"
author: "Anne Schoenauer"
date: "2025-02-13"
output: html_document
---

Note: This script requires licensed data from ecoinvent and therefore cannot be executed using the publicly available anonymized dataset. The code is provided for transparency only.
The code does not produce any figures or tables and is included solely for transparency and data quality verification.

#DESCRIPTION
## Overview of the analysis workflow
This script prepares, validates, and checks consistency of firm-level emissions and reduction potential data produced in the tilt pipeline. The workflow consists of four main steps:

### 1. Loading required libraries and setting plotting defaults
All necessary R packages for data handling, cleaning, visualization, and exporting results are loaded. A consistent plotting theme and color scales are also defined to ensure uniform visualization outputs across the analysis.

### 2. Loading input datasets
Multiple datasets are imported, including:

- Firm-level emissions results,
- Product-level company matching results,
- Activity-level emissions datasets (Ecoinvent, EPA, OECD),
- tilt activity-level data,
- Final firm-level climate action and reduction potential outputs.

Company identifiers are standardized to ensure consistent joins across datasets.

### 3. Unit tests and data quality checks
Several validation checks are applied to ensure logical consistency and data quality:

- Identification of companies with zero or missing emissions.
- Verification that Scope 1 + Scope 2 + Scope 3 emissions equal total emissions.
- Checking that mean emissions lie within the reported 5th–95th percentile interval.
- Ensuring uncertainty collapses when only one product is matched (5th–95th percentile interval = total emissions).
- Ensuring emission reduction potential does not exceed total emissions.
- Verifying that reduction components sum to total reduction potential.
- Checking consistency between Scope 2 emissions and related reduction potentials (energy related climate actions such as solar and heat pump).
- Identifying activity records missing NACE classifications.

### 4. Export of validation results
Companies or records failing validation checks are exported as CSV files for inspection and potential correction.

Overall, this script serves as a **data validation and quality assurance layer**, ensuring consistency and integrity of emissions and reduction potential estimates before further analysis or reporting.



#CODE STARTS
##01 Libraries
```{r, echo=FALSE, warning=FALSE, message=FALSE}
library(readr)
library(here)
library(dplyr)
library(readxl)
library(janitor)
library(tidyr)
library(purrr)
library(moments)
library(ggplot2)
library(scales)
library(knitr)
library(kableExtra)
library(tibble)
library(writexl)

# Set consistent plot theme
theme_set(
  theme_minimal(base_size = 12) +
    theme(
      plot.title = element_text(face = "bold", size = 14),
      plot.subtitle = element_text(size = 12),
      axis.title = element_text(face = "bold"),
      panel.grid.minor = element_blank(),
      strip.text = element_text(face = "bold", size = 12),
      legend.title = element_text(face = "bold"),
      legend.position = "bottom"
    )
)

# Optional: define viridis-based color scales (for consistency)
scale_color_viridis_d <- function(...) scale_color_viridis(discrete = TRUE, option = "D", ...)
scale_fill_viridis_d <- function(...) scale_fill_viridis(discrete = TRUE, option = "D", ...)

```

##02. Data
Load data prepped by Paul with output data
```{r, echo=FALSE, warning=FALSE, message=FALSE}

#Read in all the data
company_raw <- read_csv(here("01_input_data/01_firm_level_data/01_tilt_data/tilt_firm_level_results_anon.csv")) |> 
  mutate(
    company_id = as.character(company_id)
    )

company_product <- read_csv(here("01_input_data/02_product_level_data/tilt_related_company_product_level_results_anon.csv")) 
  
ecoinvent <- read_csv(here("01_input_data/03_activity_level_data/03_ecoinvent_data/ecoinvent_activity_level.csv")) 

epa <- read_csv(here("01_input_data/03_activity_level_data/04_epa_data/epa_activity_level.csv")) 

oecd <- read_csv(here("01_input_data/03_activity_level_data/02_oecd_data/oecd_activity_level.csv"))

lca <- read_csv(here("01_input_data/03_activity_level_data/01_tilt_data/tilt_activity_level.csv"))

##Create data to test emission reduction potential and scope 1-3 specifically
climate_actions <- company_raw %>% 
  select(c(company_id, tilt_method_s2_TE_MEAN_tonnes, reduction_potential_ab2_TE_MEAN_tonnes, reduction_potential_re1_TE_MEAN_tonnes, reduction_potential_re4_TE_MEAN_tonnes, tilt_method_reduction_potential_TE_MEAN_tonnes)) |> 
  mutate(
    company_id = as.character(company_id)
  )

```

#03. Unit tests and data quality tests
#03.01. Test if there are null total emmissions
```{r Null, echo=FALSE, warning=FALSE, message=FALSE}
#filter the ones that have tilt_method_TE_MEAN_Tonnes == 0
filtered_companies <- company_raw %>%
  filter(tilt_method_TE_MEAN_tonnes == 0) 
write_csv(filtered_companies, here("02_output_data", "01_data_quality_checks", "companies_with_NULL_TE.csv"))
```

#03.02. Test if all scope 1,2,3 add up to tilt_method_TE_MEAN_tonnes --> S1+S2+S3 = TE
```{r Scope 1-3, echo=FALSE, warning=FALSE, message=FALSE}
##tilt
# Create a new column that checks if the sum of scopes equals the total emissions
Scope_1_3 <- company_raw %>%
  mutate(
    scope_sum = tilt_method_s1_TE_MEAN_tonnes + tilt_method_s2_TE_MEAN_tonnes + tilt_method_s3_TE_MEAN_tonnes,
    is_equal = abs(scope_sum - tilt_method_TE_MEAN_tonnes) < 1e-6  # using a small tolerance
  )

# Count how many rows match
Scope_1_3_test <- Scope_1_3 %>%
  summarise(
    total_rows = n(),
    matching = sum(is_equal, na.rm = TRUE),
    not_matching = sum(!is_equal, na.rm = TRUE),
    share_matching = round(mean(is_equal, na.rm = TRUE) * 100, 2)
  )

# Filter out companies where it does NOT match
non_matching_companies <- Scope_1_3 %>%
  filter(!is_equal) %>% 
  select(company_id, tilt_method_s1_TE_MEAN_tonnes, tilt_method_s2_TE_MEAN_tonnes, tilt_method_s3_TE_MEAN_tonnes, scope_sum, tilt_method_TE_MEAN_tonnes, is_equal)

# Optional: save to CSV
write.csv(non_matching_companies, "02_output_data/01_data_quality_checks/non_matching_scope_sum.csv", row.names = FALSE)


##epa
Scope_1_3_epa <- company_raw %>%
  mutate(
    scope_sum = epa_s1_TE_MEAN_tonnes + epa_s2_TE_MEAN_tonnes + epa_s3_TE_MEAN_tonnes,
    is_equal = abs(scope_sum - epa_TE_MEAN_tonnes) < 1e-6   # using a small tolerance
  )

# Count how many rows match
Scope_1_3_test_epa <- Scope_1_3_epa %>%
  summarise(
    total_rows = n(),
    matching = sum(is_equal, na.rm = TRUE),
    not_matching = sum(!is_equal, na.rm = TRUE),
    share_matching = round(mean(is_equal, na.rm = TRUE) * 100, 2)
  )

# Filter out companies where it does NOT match
non_matching_companies_epa <- Scope_1_3_epa %>%
  filter(!is_equal) %>% 
  select(company_id,  epa_s1_TE_MEAN_tonnes, epa_s2_TE_MEAN_tonnes, epa_s3_TE_MEAN_tonnes, scope_sum, tilt_method_TE_MEAN_tonnes, is_equal)

# Optional: save to CSV
write.csv(non_matching_companies_epa, "02_output_data/01_data_quality_checks/non_matching_scope_sum_epa.csv", row.names = FALSE)

##oecd
Scope_1_3_oecd <- company_raw %>%
  mutate(
    scope_sum = oecd_s1_TE_MEAN_tonnes + oecd_s2_TE_MEAN_tonnes + oecd_s3_TE_MEAN_tonnes,
    is_equal = abs(scope_sum - oecd_TE_MEAN_tonnes) < 1e-6  # using a small tolerance
  )

# Count how many rows match
Scope_1_3_test_oecd <- Scope_1_3_oecd %>%
  summarise(
    total_rows = n(),
    matching = sum(is_equal, na.rm = TRUE),
    not_matching = sum(!is_equal, na.rm = TRUE),
    share_matching = round(mean(is_equal, na.rm = TRUE) * 100, 2)
  )

# Filter out companies where it does NOT match
non_matching_companies_oecd <- Scope_1_3_oecd %>%
  filter(!is_equal) %>% 
  select(company_id,  oecd_s1_TE_MEAN_tonnes, oecd_s2_TE_MEAN_tonnes, oecd_s3_TE_MEAN_tonnes, scope_sum, oecd_TE_MEAN_tonnes, is_equal)

# Optional: save to CSV
write.csv(non_matching_companies_oecd, "02_output_data/01_data_quality_checks/non_matching_scope_sum_oecd.csv", row.names = FALSE)
```

#03.03. Test if the mean is in between of the percentiles
```{r mean between 5th and 95th, echo=FALSE, warning=FALSE, message=FALSE}
#mean between 5th and 95th percentile
#testing it for tilt
tolerance <- 1e-6
test_interval <- company_raw %>%
  mutate(
    ci_consistent = (tilt_method_TE_p5_tonnes <= tilt_method_TE_MEAN_tonnes + tolerance) &
                    (tilt_method_TE_MEAN_tonnes <= tilt_method_TE_p95_tonnes + tolerance)
  )

filtered_companies_test_interval <- test_interval %>%
  filter(!ci_consistent) %>% 
  select(company_id, tilt_method_TE_p5_tonnes, tilt_method_TE_MEAN_tonnes, tilt_method_TE_p95_tonnes) %>% 
  filter(!(TRUE))

write.csv(filtered_companies_test_interval, "02_output_data/01_data_quality_checks/interval_tilt.csv", row.names = FALSE)


#testing it for OECD
test_interval_OECD <- company_raw %>%
  mutate(
    ci_consistent = (oecd_TE_p5_tonnes <= oecd_TE_MEAN_tonnes + tolerance) &
                    (oecd_TE_MEAN_tonnes <= oecd_TE_p95_tonnes + tolerance)
  )

filtered_companies_test_interval_oecd <- test_interval_OECD %>%
  filter(!ci_consistent) %>% 
  select(company_id, oecd_TE_p5_tonnes, oecd_TE_MEAN_tonnes, oecd_TE_p95_tonnes) %>% 
  filter(!(TRUE))

write.csv(filtered_companies_test_interval_oecd, "02_output_data/01_data_quality_checks/interval_oecd.csv", row.names = FALSE)


#testing it for EPA
test_interval_epa <- company_raw %>%
  mutate(
    ci_consistent = (epa_TE_p5_tonnes <= epa_TE_MEAN_tonnes + tolerance) &
                    (epa_TE_MEAN_tonnes <= epa_TE_p95_tonnes + tolerance)
  )

filtered_companies_test_interval_epa <- test_interval_epa %>%
  filter(!ci_consistent) %>% 
  select(company_id, epa_TE_p5_tonnes, epa_TE_MEAN_tonnes, epa_TE_p95_tonnes) %>% 
  filter(!(TRUE))

write.csv(filtered_companies_test_interval_epa, "02_output_data/01_data_quality_checks/interval_epa.csv", row.names = FALSE)
```

##03.04 Test if the 5th and 95th percentiles are the same as the mean if the product number equals 1 (no uncertainty if only 1 product)
```{r uncertainty, echo=FALSE, warning=FALSE, message=FALSE}
#If Product = 1 then no uncertainty
filtered_inconsistent_products <- company_raw %>%
  filter(number_of_products_matched == 1) %>%
  mutate(
    check_equal = tilt_method_TE_p5_tonnes == tilt_method_TE_MEAN_tonnes &
                  tilt_method_TE_MEAN_tonnes == tilt_method_TE_p95_tonnes
  ) %>%
  filter(!check_equal)

write_csv(filtered_inconsistent_products, here("02_output_data", "01_data_quality_checks", "inconsistent_products.csv"))
```


#03.05. Testing if the emission reduction potential is <= Total emissions
```{r emission_reduction_potential small/equal than total emissions, echo=FALSE, warning=FALSE, message=FALSE}
filtered_companies_fail_ERP_test <-
  climate_actions %>%
  select(
    company_id,
    tilt_method_reduction_potential_TE_MEAN_tonnes
  ) %>%
  left_join(
    company_raw %>%
      select(
        company_id,
        tilt_method_TE_MEAN_tonnes
      ),
    by = "company_id"
  ) %>%
  mutate(
    reduction_logical =
      tilt_method_reduction_potential_TE_MEAN_tonnes <=
      tilt_method_TE_MEAN_tonnes
  ) %>%
  filter(!reduction_logical)

write.csv(filtered_companies_fail_ERP_test, "02_output_data/01_data_quality_checks/emission_reduction_potential_bigger_than_emissions.csv", row.names = FALSE)
```

#03.06. Testing if all climate actions (AB2 = efficient machinery, RE1 = solar panels, RE4= Heat pumps) equal total emission reduction potentail 
```{r emission_reduction_potential, echo=FALSE, warning=FALSE, message=FALSE}
#AB2 + RE1 + RE4 = ERP
sum_ERP_test <- company_raw %>%
  mutate(
    reduction_potential_sum = reduction_potential_ab2_TE_MEAN_tonnes +
                              reduction_potential_re1_TE_MEAN_tonnes +
                              reduction_potential_re4_TE_MEAN_tonnes,

    reduction_consistent = abs(reduction_potential_sum - tilt_method_reduction_potential_TE_MEAN_tonnes) < 1e-6
  )

filtered_companies_fail_sum_ERP_test <- sum_ERP_test %>%
  filter(!reduction_consistent) %>% 
  select(company_id, reduction_potential_ab2_TE_MEAN_tonnes, reduction_potential_re1_TE_MEAN_tonnes, reduction_potential_re4_TE_MEAN_tonnes,
         tilt_method_reduction_potential_TE_MEAN_tonnes) %>% 
  mutate(across(where(is.numeric), ~ format(.x, scientific = FALSE)))

write_csv(filtered_companies_fail_sum_ERP_test, here("02_output_data", "01_data_quality_checks", "companies_with_sum_of_climate_action_unequal_total_reduction_potential.csv"))
```


#03.07.  Test if the energy related climate actions (RE1 and RE4) are not higher than scope 2
```{r emission_reduction_potential, echo=FALSE, warning=FALSE, message=FALSE}
##Filter out reduction_potential_ab2_TE_MEAN_tonnes > 0 -->
sum_S2_test <- climate_actions %>% 
  filter(!(reduction_potential_ab2_TE_MEAN_tonnes > 0)) %>% 
  mutate(
    inconsistent =  
      tilt_method_reduction_potential_TE_MEAN_tonnes >
      tilt_method_s2_TE_MEAN_tonnes + tolerance
  ) %>% 
  filter(inconsistent == TRUE) %>% 
  select(company_id, reduction_potential_ab2_TE_MEAN_tonnes, reduction_potential_re1_TE_MEAN_tonnes, reduction_potential_re4_TE_MEAN_tonnes,
         tilt_method_reduction_potential_TE_MEAN_tonnes, tilt_method_s2_TE_MEAN_tonnes) %>% 
  mutate(across(where(is.numeric), ~ format(.x, scientific = FALSE)))

write_csv(sum_S2_test, here("02_output_data", "01_data_quality_checks", "sum_s2_inconsistent.csv"))
```


##03.08 Test if all LCA/sectors have a NACE code
```{r NACE codes, echo=FALSE, warning=FALSE, message=FALSE}
##Ecoinvent no NACE code
ecoinvent_NA <- ecoinvent%>% 
  filter(nace_2_list == "[]")

write_csv(ecoinvent_NA, here("02_output_data", "01_data_quality_checks","ecoinvent_with_no_NACE_code.csv"))

##EPA no NACE code
epa_NA <- epa %>% 
  filter(nace_2_list == "[]")

write_csv(ecoinvent_NA, here("02_output_data","01_data_quality_checks", "epa_with_no_NACE_code.csv"))

##OECD no NACE code
oecd_NA <- oecd %>% 
  filter(nace_2 == "[]")

write_csv(ecoinvent_NA, here("02_output_data", "01_data_quality_checks","oecd_with_no_NACE_code.csv"))
```
