---
title: "02b_stratification"
author: "Tilman Trompke"
date: "2025-02-13"
output: html_document
---

The code generates Table 4 "Overview table of sample set counts and shares as well as population size counts and shares and weights per stratum, incl. shares (in%) compared to total dataset" in the Appendix. 

## Overview of the analysis workflow
This R Markdown script processes German SME population benchmarks and firm-level sample data, constructs common strata by NACE section and employee size class, evaluates sample coverage across strata, and derives post-stratification weights.

### 01. Loading in the Libraries
Loads the R packages needed for reading files, cleaning/wrangling data, and producing tables.

### 02. Preparing the Data
Loads and prepares two datasets:
  - Population benchmarks (Germany SME factsheet Excel): cleans messy Excel structure (merged cells/blank rows), keeps 
  - SME size classes (micro/small/medium), and extracts NACE identifiers (section letter, 2-digit code).
Sample firm-level data (CSV): filters to SMEs (employees < 250, revenue ≤ 50m) and keeps only firms with positive emissions; creates the same micro/small/medium size classes.

## 03. Population vs. sample data
Aligns both datasets to the same strata (NACE section × employee size class), then:
  - counts firms per stratum in the sample (n_j)
  - totals firms per stratum in the population (N_j)
  - computes post-stratification weights per stratum: weight = N_j / n_j
  - reshapes the results into a wide summary table (micro/small/medium columns + totals)
  - formats counts as N (percent%) for display
  - creates a LaTeX table and writes it out as a .tex file for inclusion in a report/paper


##01. Libraries
```{r}

library(readr)
library(here)
library(dplyr)
library(janitor)
library(readxl)
library(tidyr)
library(stringr)
library(knitr)
library(kableExtra)
library(purrr)


```

##02. Data
#02.1. Load data
Germany factsheet is loaded from this page: https://single-market-economy.ec.europa.eu/smes/sme-strategy-and-sme-friendly-business-conditions/sme-performance-review_en#sba-fact-sheets 
```{r, echo=FALSE, warning=FALSE, message=FALSE}
# load population data

pop_raw <- read_xlsx(here("01_input_data/01_firm_level_data/03_Population_Germany/Germany_SME_key_figures_2025.xlsx"), 
                  sheet ="NoE 2DGT", 
                  skip = 8, # remove first 8 rows because they have no values
                  .name_repair = "minimal") |> # avoid that r automatically renames column names if duplicates exist because I remove duplicate columns in next step
  select(1:20) |> # remove irrelevant columns with same name
  clean_names() 
  

# load sample tilt-data
sample_raw <- read_csv(here("01_input_data/01_firm_level_data/01_tilt_data/tilt_firm_level_results_anon.csv")) |> 
  mutate(company_id = as.character(company_id),  
  nace_section_list_adjusted = gsub('[\\[\\]"]', '', nace_section_list_adjusted)) |>
  filter(EmployeesTotal < 250) |> 
  filter(Revenue <= 50000000) |> 
  filter(tilt_method_TE_MEAN_tonnes > 0) |> 
  rename(
    nace_section_list = nace_section_list_adjusted
  )

```

#02.2. Clean population data - Germany SME key Figurs 2025
```{r, echo=FALSE, warning=FALSE, message=FALSE}

colnames(pop_raw)

# remove empty row
pop_raw <- pop_raw[-1,]

# solve merged cell problem --> many empty values in left column
pop_raw <- pop_raw |>
  mutate(division_nace_rev_2 = na_if(division_nace_rev_2, "")) %>%  # treat "" as missing
  fill(division_nace_rev_2, .direction = "down")

# fill in NAs in 2025 columns with next best year
pop_raw <- pop_raw |>
  rowwise() |>
  mutate(x2025 = coalesce(x2025, x2024, x2023, x2022, x2021, x2020, x2019, x2018)) |>
  ungroup()

# subselection
  # remove whitespaces etc
  pop <- pop_raw |>
    mutate(size_class = str_squish(size_class))
  
  # remove irrelevant size classes
  pop <- pop %>%
    filter(size_class %in% c("0 - 9", "10 - 49", "50 - 249"))

  # sanity check
  table(pop$size_class)

  # select relevant columns
  pop <- pop |> 
    select(division_nace_rev_2, size_class, x2025)
  
# create additional column with nace 2 digit
  pop <- pop |>
  mutate(
    x2_nace = str_extract(division_nace_rev_2, "(?<=^[A-Z])\\d{2}")
  ) |>
  relocate(x2_nace, .after = division_nace_rev_2)

  # sanity check
  unique(pop$x2_nace)
  
# create additional column with nace section
pop <- pop |>
  mutate(
    section_nace = str_extract(division_nace_rev_2, "^[A-Z]")
  ) |>
  relocate(section_nace, .after = division_nace_rev_2)

  # sanity check
  unique(pop$section_nace)
  
# create size class name 
  pop <- pop |> 
  mutate(emp_size_class = case_when(
    size_class == "0 - 9" ~ "micro", 
    size_class == "10 - 49" ~ "small", 
    size_class == "50 - 249" ~ "medium",
    TRUE ~ NA_character_
  )) |> 
  relocate(emp_size_class, .after = size_class)

rm(pop_raw)
  
```

#02.3. Clean sample data
# # select relevant columns
```{r, echo=FALSE, warning=FALSE, message=FALSE}
sample <- sample_raw |> 
  select(
    company_id, 
    nace_section, 
    nace_section_list, 
    EmployeesTotal
  )

```


#02.4. Create Size Class for stratification
```{r, echo=FALSE, warning=FALSE, message=FALSE}
# create emp size class label (sector labels already exists)
sample <- sample |>
  mutate(
    emp_size_class = case_when(
      EmployeesTotal >= 0   & EmployeesTotal <= 9   ~ "micro",
      EmployeesTotal >= 10  & EmployeesTotal <= 49  ~ "small",
      EmployeesTotal >= 50  & EmployeesTotal <= 249 ~ "medium",
      TRUE ~ NA_character_
    ))

  # sanity check
  table(sample$emp_size_class)

```


##03. Population vs. sample data
#03.1. Create same strata (nace Section x employee size class) in both the population dataset and the sample dataset
```{r, echo=FALSE, warning=FALSE, message=FALSE}
## SAMPLE
# create stratum (nace section x employee size class) column  in sample data
  sample <- sample |>
  mutate(stratum = paste(nace_section, emp_size_class, sep = "_")) 

## POPULATION
# create stratum (nace section x employee size class) column in pop data
  pop <- pop |>
  mutate(stratum = paste(section_nace, emp_size_class, sep = "_")) |> 
    relocate(stratum, .after = emp_size_class)

# sanity check
table(pop$stratum)
table(sample$stratum)
  
```

#03.2. Compute population / sample counts
```{r, echo=FALSE, warning=FALSE, message=FALSE}
## SAMPLE 
sample_counts <- sample |>
  group_by(stratum) |>
  summarise(
    n_j = n(),
    .groups = "drop"
  )

  summary(sample_counts)
  
  # count share of strata below 20 
  sample_counts %>%
  summarise(share_below_20 = mean(n_j < 20))
  
  # count share of strata below 25
  sample_counts %>%
  summarise(share_below_20 = mean(n_j < 25))

  
## POPULATION
pop_counts <- pop |>
  group_by(stratum) |>
  summarise(
    N_j = sum(x2025),   # or whatever the population count variable is
    .groups = "drop"
  )

  summary(pop_counts)
  
  # count share of strata below 25
  pop_counts %>%
  summarise(share_below_20 = mean(N_j < 25))

# Overview  
sample_counts
pop_counts

```

#03.3. Create weights
```{r, echo=FALSE, warning=FALSE, message=FALSE}

# create weights
weights_df <- sample_counts %>%
  left_join(pop_counts, by = "stratum")

summary(weights_df$N_j)
summary(weights_df$n_j)

# calculate weights
weights_df <- weights_df |>
  mutate(weight = N_j / n_j)

```

#03.4. Create summary weights table with sample size per stratum for sample, pop, and weights
```{r, echo=FALSE, warning=FALSE, message=FALSE}

weights_df_wide <- weights_df |> 
  mutate(weight = round(weight, 2)) |> 
  separate(
    stratum,
    into = c("sector", "size"),
    sep = "_",
    remove = TRUE
  ) |>
  mutate(size = factor(size, levels = c("micro", "small", "medium"))) |>
  pivot_longer(
    cols = c(n_j, N_j, weight),
    names_to = "measure",
    values_to = "value"
  ) |>
  pivot_wider(
    id_cols = sector,
    names_from = c(measure, size),
    values_from = value,
    names_sep = "_"
  ) |>
  arrange(sector) |> 
  select(sector, n_j_micro, n_j_small, n_j_medium, N_j_micro, N_j_small, N_j_medium, weight_micro, weight_small, weight_medium)

weights_df_wide <- weights_df_wide |>
  mutate(
    n_j_total    = rowSums(across(all_of(c("n_j_micro","n_j_small","n_j_medium"))), na.rm = TRUE),
    N_j_total    = rowSums(across(all_of(c("N_j_micro","N_j_small","N_j_medium"))), na.rm = TRUE),
    weight_total = N_j_total / n_j_total
  )

weights_df_wide <- weights_df_wide |>
  bind_rows(
    weights_df_wide |>
      summarise(
        across(
          where(is.numeric),
          ~ sum(.x, na.rm = TRUE)
        )
      ) |>
      mutate(sector = "Total")
  ) |> 
    select(sector, n_j_micro, n_j_small, n_j_medium, n_j_total, N_j_micro, N_j_small, 
           N_j_medium, N_j_total, weight_micro, weight_small, weight_medium, weight_total)


```

#03.5. Create pretty table with %-values
```{r, echo=FALSE, warning=FALSE, message=FALSE}

n_j_cols <- c("n_j_micro", "n_j_small", "n_j_medium", "n_j_total")
N_j_cols <- c("N_j_micro", "N_j_small", "N_j_medium", "N_j_total")

n_j_grand <- weights_df_wide |> 
  filter(sector == "Total") |> 
  summarise(x = first(n_j_total)) |> 
  pull(x)

N_j_grand <- weights_df_wide |> 
  filter(sector == "Total") |> 
  summarise(x = first(N_j_total)) |> 
  pull(x)

weights_df_wide_display <- weights_df_wide |>
  mutate(
    across(
      all_of(n_j_cols),
      ~ sprintf(
        "%d (%.1f%%)",
        as.integer(.x),
        if (!is.na(n_j_grand) && n_j_grand > 0) 100 * as.numeric(.x) / n_j_grand else NA_real_
      )
    ),
    across(
      all_of(N_j_cols),
      ~ sprintf(
        "%d (%.1f%%)",
        as.integer(.x),
        if (!is.na(N_j_grand) && N_j_grand > 0) 100 * as.numeric(.x) / N_j_grand else NA_real_
      )
    )
  )


```

#03.6 Latex format & export Table 4
```{r, echo=FALSE, warning=FALSE, message=FALSE}
latex_tbl_counts <- weights_df_wide_display |>
  kable(
    format = "latex",
    booktabs = TRUE,
    escape = TRUE,
    align = "l"
  ) |>
  add_header_above(
    c(
      " " = 1,
      "sample counts per stratum" = 4,
      "population counts per stratum" = 4,
      "weights per stratum" = 4
    )
  ) |>
  kable_styling(
    latex_options = c(
      "hold_position",
      "scale_down"      # auto-scale to page width
    ),
    font_size = 8       # fits A4 landscape well
  ) |>
  column_spec(1, width = "2.5cm") |>  # sector column
  column_spec(2:13, width = "1.6cm")  # all numeric columns

latex_tbl_counts

writeLines(latex_tbl_counts, here("02_output_data/02_stratification/Table-4_sample_pop_weights_per_stratum.tex"))
```

##04. Attachement to firm data
We assign the strata weights to the firm list for later usage in the analysis section
```{r}
# attach weights
  sample_weights <- sample |>
    left_join(
      weights_df %>% select(stratum, weight),
      by = "stratum"
    )

```


##07. Save data
```{r}
write_csv(sample_weights, here("02_output_data/02_stratification/sample_with_weights.csv")) # used in analysis section
write_csv(weights_df_wide_display, here("02_output_data/02_stratification/weights_per_stratum_percentages.csv")) # shown in Supplementary Material
```

