---
title: "03_Matching_Algorithm_checks"
author: "Anne Schoenauer"
date: "2025-02-13"
output: html_document
---

Note: This script requires licensed data from ecoinvent and therefore cannot be executed using the publicly available anonymized dataset. The code is provided for transparency only.

The code generates Table 5 "Evaluation Metrics for the multi-agent LCA matching system" in the Appendix. 

## Overview of the analysis workflow
Overview of the matching algorithm evaluation workflow. This script evaluates the performance of the activity-level matching algorithm using manual validation as ground truth. The workflow focuses on assessing precision, error, latency, and coverage of the matching results.

### 1. Loading required libraries and setting plotting defaults
All required R packages for data manipulation, evaluation, visualization, and table export are loaded. A consistent ggplot2 theme and color scales are defined to ensure uniform presentation of results.

### 2. Loading and preparing validation datasets
Multiple datasets are imported and combined, including:
  - Activity-level matching validation data
  - Model-selected matches after filtering 
  - Activity-level environmental footprint data 
  - Firm- and product-level pipeline outputs
Datasets are joined using company, product, and activity identifiers to align model predictions with manually accepted matches.

### 3. Evaluation of matching quality
The script compares model-selected activities to manually accepted candidates to assess matching accuracy:
  - Precision@1: share of cases where the model-selected match equals the accepted match
  - Precision@10: share of cases where an accepted match appears in the evaluated candidate list
  - MAPE: mean absolute percentage error of environmental footprint values for incorrectly matched cases
  - Numeric values are standardized prior to error computation to ensure consistency.

### 4. Latency and score assessment
For matches selected after filtering, the script computes average:
  - Matching evaluation scores
  - Runtime (in seconds)
to assess computational performance and model confidence.

### 5. Coverage analysis
Coverage is evaluated by comparing extracted products to successfully matched products at the firm level. Summary statistics quantify:
  - Firms with no extracted products
  - Total extracted versus matched products
  - The share of missing matches

### 6. Export of evaluation results
All evaluation metrics are displayed in the HTML output and exported as LaTeX tables for integration into reports and publications.


##01. Libraries
```{r, echo=FALSE, warning=FALSE, message=FALSE}
library(readr)
library(here)
library(dplyr)
library(stringr)
library(ggplot2)
library(scales)
library(stringr)
library(purrr)
library(latex2exp)
library(knitr)
library(kableExtra)
library(viridis)
library(ggplot2)
library(showtext)

# Set consistent plot theme
theme_set(
  theme_minimal(base_size = 12) +
    theme(
      plot.title = element_text(face = "bold", size = 14),
      plot.subtitle = element_text(size = 12),
      axis.title = element_text(face = "bold"),
      panel.grid.minor = element_blank(),
      strip.text = element_text(face = "bold", size = 12),
      legend.title = element_text(face = "bold"),
      legend.position = "bottom"
    )
)

# Optional: define viridis-based color scales (for consistency)
scale_color_viridis_d <- function(...) scale_color_viridis(discrete = TRUE, option = "D", ...)
scale_fill_viridis_d <- function(...) scale_fill_viridis(discrete = TRUE, option = "D", ...)

```



##02. Creating the first part of the evaluation metric (MAPE, precision@1 and precision@10)
```{r, warning=FALSE, message=FALSE}
# For false positives, precision.
Matching_validation <- read_csv2(
  here("01_input_data", "03_activity_level_data", "01_tilt_data", "Matching_validation.csv")
) %>%
  select(
    company_id,
    lca_id,
    related_company_product,
    total_emissions_per_eur,
    activity_name,
    match_eval_score,
    match_selected_after_filtering,
    validation_Anne
  ) %>%
  distinct(company_id, lca_id, activity_name, related_company_product, .keep_all = TRUE) %>%
  rename(co2_footprint_per_euro = total_emissions_per_eur) %>%
  filter(co2_footprint_per_euro > 0)

#Extract the product predictions
predicted <- Matching_validation %>%
  filter(match_selected_after_filtering == TRUE) %>%
  select(
    company_id,
    related_company_product,
    predicted_EF = co2_footprint_per_euro,
    validation_Anne
  )

#Extract the ground truth
truth <- Matching_validation %>%
  filter(validation_Anne == 1) %>%
  select(
    company_id,
    related_company_product,
    true_EF = co2_footprint_per_euro
  )

false_matches <- predicted %>%
  filter(validation_Anne == 0)

#Combine the predictions and truth
mape_data <- false_matches %>%
  left_join(
    truth,
    by = c("company_id", "related_company_product")
  ) %>%
  mutate(
    ape = abs(predicted_EF - true_EF) / abs(true_EF)
  )

mape <- mean(mape_data$ape, na.rm = TRUE) * 100


# ----------------------------
# compute Precision@1 (product-level)
# ----------------------------
precision_at_1 <- Matching_validation %>%
  filter(match_selected_after_filtering == TRUE) %>%
  summarise(precision_at_1 = mean(validation_Anne == 1, na.rm = TRUE)) %>%
  pull(precision_at_1)

cat("Precision@1:", round(precision_at_1, 3), "\n")

# ----------------------------
# compute Precision@K (our "list" size, using top K candidates by match_eval_score)
# If the candidate list is shorter than K, we only use all available candidates.
# ----------------------------
precision10_data <- Matching_validation %>%
  
  # rank candidates within each product
  group_by(company_id, related_company_product) %>%
  arrange(desc(match_eval_score), .by_group = TRUE) %>%
  
  # assign rank
  mutate(rank = row_number()) %>%
  
  # keep top 10 (or fewer if less exist)
  filter(rank <= 10) %>%
  
  # check if a true match is inside top 10
  summarise(
    hit = any(validation_Anne == 1, na.rm = TRUE),
    .groups = "drop"
  )

precision_at_10 <- mean(precision10_data$hit, na.rm = TRUE)


## Export the metrics table
evaluation_table <- tibble(
  Metric = c("Precision@1", paste0("Precision@"), "MAPE"),
  Value = c(round(precision_at_1, 3),
            round(precision_at_10, 3),
            round(mape, 2))
)

latex_evaluation_table <- kable(
  evaluation_table,
  format = "latex",
  booktabs = TRUE,
  caption = "Evaluation Metrics"
)

writeLines(
  latex_evaluation_table,
  here("02_output_data/03_matching_algorithm_validation/Table-5-first-part_evaluation_table.tex")
)

# Output
evaluation_table
```



#0.3 For latency score average
```{r, warning=FALSE, message=FALSE}
Latency_matching_score <- read_delim(
  here("01_input_data/03_activity_level_data/01_tilt_data/tilt_activity_level.csv"),
  delim = ",", 
  locale = locale(encoding = "UTF-8"),
  show_col_types = FALSE
)

summary_stats_latency <- Latency_matching_score %>%
  filter(match_selected_after_filtering == TRUE) %>% 
  summarise(
    avg_matching_score = mean(match_eval_score, na.rm = TRUE),
    avg_run_time = mean(runtime_seconds, na.rm = TRUE)
    )

latex_summary_stats_latency <- kable(
  summary_stats_latency,
  format = "latex",
  booktabs = TRUE,
  caption = "Latency Stats"
)

summary_stats_latency
writeLines(latex_summary_stats_latency, here("02_output_data/03_matching_algorithm_validation/Table-5-second-part-latency_score_table.tex"))

```


#0.4 For coverage
```{r, warning=FALSE, message=FALSE}
# Read and prepare the data
Coverage <- read_csv(here("01_input_data/02_product_level_data/tilt_related_company_product_level_results_anon.csv")) %>% 
  select(company_id, number_of_products_matched, number_of_products_extracted) %>%
  mutate(
    number_of_extracted_products = as.numeric(number_of_products_extracted),
    number_of_matched_products = as.numeric(number_of_products_matched)
  )

# Collapse to one row per company
Coverage_firm_level <- Coverage %>%
  group_by(company_id) %>%
  summarise(
    number_of_extracted_products = sum(number_of_extracted_products, na.rm = TRUE),
    number_of_matched_products = sum(number_of_matched_products, na.rm = TRUE),
    .groups = "drop"
  )

# Calculate summary statistics
summary_stats_coverage <- Coverage_firm_level %>%
  summarise(
    n_companies_no_extraction = sum(number_of_extracted_products == 0, na.rm = TRUE),
    total_extracted_products = sum(number_of_extracted_products, na.rm = TRUE),
    total_matched_products = sum(number_of_matched_products, na.rm = TRUE),
    total_missing_matches = total_extracted_products - total_matched_products,
    percent_missing_matches = total_missing_matches / total_extracted_products * 100,
    total_firms = n(),
    percent_firms_no_extraction = n_companies_no_extraction / total_firms * 100
  )

#Output
latex_summary_stats_coverage <- kable(
  summary_stats_coverage,
  format = "latex",
  booktabs = TRUE,
  caption = "Coverage"
)


writeLines(latex_summary_stats_coverage, here("02_output_data/03_matching_algorithm_validation/Table-5-third-part_coverage.tex"))


```
