---
title: "06a_analysis_sectoral_emission"
author: "Tilman Trompke and Anne Schoenauer"
date: "2025-02-13"
output: html_document
---

##Overview of the analysis workflow
This R Markdown analyzes sectoral emissions and reduction potentials based on SME firm-level results from the tilt pipeline. It combines weighted sector aggregates (using stratification weights), geographic diagnostics (Germany map using proprietary coordinates), and sector/industry deep-dives that relate emissions intensity to reduction potential.

###01. Libraries
All required R packages for data processing, plotting, LaTeX support, and spatial mapping are loaded. A consistent ggplot theme and viridis-based helper scales are defined for uniform visualization outputs.

###02. Preparing the data
Firm-level tilt results are loaded for SMEs (Employees < 250; Revenue ≤ 50m) and enriched with (i) sample weights for population-representative aggregation and (ii) a company-id to latitude/longitude mapper for geographic visualization. OECD sector totals are also loaded for benchmarking. The script prepares additional variables used throughout the analysis, including sector labels, revenue classes, reduction-potential shares, and emission-intensity measures (kg CO₂e per EUR revenue). Two industry subsets are created for deep-dives: A01 (agriculture) and C20 (chemicals).

###03. Geographic distribution of firms
Using proprietary longitude/latitude data, the script maps firms within Germany and visualizes their emissions rank (normalized). This section generates Figure 4 (code currently set to eval = FALSE, so it will only render if enabled) and saves it as:
Figure 4: 02_output_data/06_analysis/te_ger_heatmap.pdf

##04. Analysis results
This part produces the main sector-level outputs:
- Figure 5: Compares sectoral shares of total emissions between OECD economy-wide totals and weighted tilt totals, focusing on the top 5 sectors by tilt share and grouping the remainder as “Other Sectors”.
Output: 02_output_data/06_analysis/Figure-6-oecd_tilt_plot_share_emissions_top5_labeled.pdf
- Figure 6: Creates a sector-by-climate-action heatmap of emission reduction potential (in kt CO₂e, with percent shares relative to sector emissions), excluding selected sectors and focusing on applicable actions.
Output: 02_output_data/06_analysis/reduction_potential_heatmap.pdf
- Figure 7: Deep-dive for A01 (agriculture) showing a quadrant/bubble plot of emissions intensity vs. reduction-potential intensity, with median-based quadrant cutoffs and an annotation for the upper-right quadrant (high emissions & high reduction potential).
Output: 02_output_data/06_analysis/quadrants_a01.pdf
- Figure 13 (Appendix 9.4): аналогous quadrant/bubble deep-dive for C20 (chemicals).
Output: 02_output_data/06_analysis/quadrant_c20.pdf

Overall, this script provides an economy- and sector-level view of emissions and mitigation opportunities, combining population-weighted aggregation with geographic diagnostics and industry-specific deep-dives into emissions intensity and reduction potential.


##01. Libraries
```{r, echo=FALSE, warning=FALSE, message=FALSE}
# Load all required libraries
library(readr)
library(here)
library(dplyr)
library(stringr)
library(ggplot2)
library(scales)
library(ggforce)
library(latex2exp)
library(knitr)
library(kableExtra)
library(viridis)
library(ggplot2)
library(showtext)
library(readxl)
library(here)
library(tidyr)
library(dplyr)
library(ggplot2)
library(sf)
library(scales)
library(rnaturalearth)
library(rnaturalearthdata)



# Set consistent plot theme
theme_set(
  theme_minimal(base_size = 12) +
    theme(
      plot.title = element_text(face = "bold", size = 14),
      plot.subtitle = element_text(size = 12),
      axis.title = element_text(face = "bold"),
      panel.grid.minor = element_blank(),
      strip.text = element_text(face = "bold", size = 12),
      legend.title = element_text(face = "bold"),
      legend.position = "bottom"
    )
)

# Use a color scale from viridis (for consistency and accessibility)
scale_color_viridis_d <- function(...) scale_color_viridis(discrete = TRUE, option = "D", ...)
scale_fill_viridis_d <- function(...) scale_fill_viridis(discrete = TRUE, option = "D", ...)

```

##02. Preparing the Data
#02.1. Loading in data
We load in firm-level data and select total emissions (mean, 95th and 5th uncertainty range). We also create a SD. 
```{r, warning=FALSE, message=FALSE}
#firm-level data
TE_tilt_raw <- read_csv(here("01_input_data/01_firm_level_data/01_tilt_data/tilt_firm_level_results_anon.csv")) %>% 
  filter(EmployeesTotal <250) %>% 
  filter(Revenue <= 50000000) 

#sample weights
sample_weights <- read_csv(here("02_output_data/02_stratification/sample_with_weights.csv"))

#proprietary longitude and latitude
long_lat_mapper <- read_csv(here("01_input_data/01_firm_level_data/01_tilt_data/mapper_company_id_longitude_latitude.csv")) 

# oecd data
oecd_emissions_raw <- read_csv(here("01_input_data/03_activity_level_data/02_oecd_data/oecd_emission_factors.csv"))

```


#02.2. Prep company data
```{r}
#Cleaning of NACE data and creation of sector_label 
TE_tilt <- TE_tilt_raw %>%
   select(c(company_id, Revenue, EmployeesTotal, nace_section_list_adjusted, nace_section, nace_2_list, tilt_method_lca_source,
          tilt_method_TE_MEAN_tonnes, tilt_method_TE_SD_tonnes, tilt_method_TE_p95_tonnes, 
          tilt_method_TE_p5_tonnes, tilt_method_s1_TE_MEAN_tonnes, tilt_method_s1_TE_p5_tonnes, tilt_method_s1_TE_p95_tonnes,tilt_method_s2_TE_MEAN_tonnes, tilt_method_s2_TE_p5_tonnes, tilt_method_s2_TE_p95_tonnes, tilt_method_s3_TE_MEAN_tonnes, tilt_method_s3_TE_p5_tonnes, tilt_method_s3_TE_p95_tonnes, 
          tilt_method_reduction_potential_TE_MEAN_tonnes, 
          reduction_potential_ab2_TE_MEAN_tonnes, 
          reduction_potential_re1_TE_MEAN_tonnes, 
          reduction_potential_re4_TE_MEAN_tonnes, number_of_products_matched))


# make sector-label "multi_sector" if multiple sectors exist, or use the nace name
TE_tilt <- TE_tilt %>%
  mutate(
    sector_clean = ifelse(nace_section == "multi-sector", "multi-sector", nace_section_list_adjusted)
  )

# create reduction potential share
TE_tilt <- TE_tilt |> 
  mutate(reduction_potential_share =
           tilt_method_reduction_potential_TE_MEAN_tonnes/tilt_method_TE_MEAN_tonnes)

sample_weights <- sample_weights %>% 
  select(c(stratum, weight, emp_size_class, company_id))

#0Join weights to firm data
TE_tilt <- TE_tilt |> 
  left_join(sample_weights, by = "company_id")

#########


##Create the data for German map: Join for geographic distribution also the Latitude/Longitude (Propietary data)
TE_tilt_long_lat <- TE_tilt %>% 
  left_join(long_lat_mapper, by ="company_id")


#########

##Create the data for Deep-dive: 
#Select relevant columns and calculate emission factors
TE_tilt_deep_dive <- TE_tilt_raw |> 
  select(company_id, nace_section,
         EmployeesTotal, Revenue, tilt_method_s1_TE_MEAN_tonnes, tilt_method_reduction_potential_TE_MEAN_tonnes, nace_2_list, tilt_method_s2_TE_MEAN_tonnes, tilt_method_s3_TE_MEAN_tonnes, oecd_s1_TE_MEAN_tonnes, oecd_s2_TE_MEAN_tonnes, oecd_s3_TE_MEAN_tonnes, epa_s1_TE_MEAN_tonnes, epa_s2_TE_MEAN_tonnes, epa_s3_TE_MEAN_tonnes) |> 
  mutate(
    tilt_method_s1_EF_MEAN_kg = (tilt_method_s1_TE_MEAN_tonnes/Revenue)*1000, 
    tilt_method_s2_EF_MEAN_kg = (tilt_method_s2_TE_MEAN_tonnes/Revenue)*1000, 
    tilt_method_s3_EF_MEAN_kg = (tilt_method_s3_TE_MEAN_tonnes/Revenue)*1000, 
    tilt_method_EF_MEAN_kg = (tilt_method_s1_EF_MEAN_kg+tilt_method_s2_EF_MEAN_kg+tilt_method_s3_EF_MEAN_kg),
    oecd_s1_EF_MEAN_kg = (oecd_s1_TE_MEAN_tonnes/Revenue)*1000, 
    oecd_s2_EF_MEAN_kg = (oecd_s2_TE_MEAN_tonnes/Revenue)*1000, 
    oecd_s3_EF_MEAN_kg = (oecd_s3_TE_MEAN_tonnes/Revenue)*1000, 
    oecd_EF_MEAN_kg = (oecd_s1_EF_MEAN_kg+oecd_s2_EF_MEAN_kg+oecd_s3_EF_MEAN_kg),
    epa_s1_EF_MEAN_kg = (epa_s1_TE_MEAN_tonnes/Revenue)*1000, 
    epa_s2_EF_MEAN_kg = (epa_s2_TE_MEAN_tonnes/Revenue)*1000, 
    epa_s3_EF_MEAN_kg = (epa_s3_TE_MEAN_tonnes/Revenue)*1000,
    epa_EF_MEAN_kg = (epa_s1_EF_MEAN_kg+epa_s2_EF_MEAN_kg+epa_s3_EF_MEAN_kg), 
    tilt_method_reduction_potential_EF_MEAN_kg = (tilt_method_reduction_potential_TE_MEAN_tonnes/Revenue *1000)
  )

#Create subset of data
# select single sections
TE_tilt_single <- TE_tilt_deep_dive |> 
  filter(!(nace_section == "multi-sector"))


#Create relevant subsets for deep-dive analysis
#For orientation: List of NACE codes: https://nacecode.de/ or https://ec.europa.eu/eurostat/documents/3859598/5902521/KS-RA-07-015-EN.PDF.pdf/dd5443f5-b886-40e4-920d-9df03590ff91?t=1414781457000 
#Illustrative selection (all other NACE codes also possible): 
#- A01: Crop and animal production, hunting and related service activities 
#- C20: Manufacture of chemicals and chemical products 

tilt_a01 <- TE_tilt_single |> 
  filter(nace_2_list == "A01")

tilt_c20 <- TE_tilt_single |> 
  filter(nace_2_list == "C20")


#########

##Create data for OECD vs. tilt comparison
#weight total emissions using the weights of the stratification exercise
TE_tilt_weighted <- TE_tilt |> 
  mutate(
    total_emissions_weighted = tilt_method_TE_MEAN_tonnes*weight,
    total_revenue_weighted = Revenue*weight, 
    total_s1_weighted = tilt_method_s1_TE_MEAN_tonnes*weight, 
    total_s2_weighted = tilt_method_s2_TE_MEAN_tonnes*weight, 
    total_s3_weighted = tilt_method_s3_TE_MEAN_tonnes*weight
  ) 


#Sum emissions per sector letter
company_emissions_by_sector <- TE_tilt_weighted %>%
  filter(!(nace_section == "multi-sector")) |> 
  group_by(nace_section) |> 
  summarise(
    total_company_emissions = sum(total_emissions_weighted, na.rm = TRUE),
    total_company_s1 = sum(total_s1_weighted, na.rm = TRUE), 
    total_company_s2 = sum(total_s2_weighted, na.rm = TRUE), 
    total_company_s3 = sum(total_s3_weighted, na.rm = TRUE), 
    .groups = "drop"
  ) |> 
  mutate(
    nace_section = as.character(nace_section)
  )

oecd_emissions <- oecd_emissions_raw |>
  mutate(
    nace_section = str_extract(activity, "[A-Z]")
  ) |> 
  select(origin_area, activity, time_period, S1, S2, S3U, total_s1_s2_s3u, prod_value, nace_section)

# note: OECD emissions are displayed in million kg, but we want to show them in tonnes, i.e. divide by 1,000 (conversion to tons) and multiply with 1,000,000 (from millions to normal count)
# multiply emissions & prod with 1,000,000 (because they are in million kg)
oecd_emissions <- oecd_emissions |> 
  mutate(
    total_s1_s2_s3u = total_s1_s2_s3u * 1000000 / 1000, 
    S1 = S1* 1000000 / 1000, 
    S2 = S2* 1000000 / 1000, 
    S3U = S3U* 1000000 / 1000
  )



```


## 03. Geographic Distribution of Firms
This generates Figure 4: SMEs locations coloured by emission rank, unweighted; with propietary data. 

```{r map-distribution, echo=FALSE, message=FALSE, warning=FALSE}

##Create Rank (firm with the lowest emissions get 1 the next lowest 2, etc, we normalise the rank so that the rank is between 0-1)
TE_tilt_long_lat <- TE_tilt_long_lat %>%
  mutate(
    tilt_method_TE_MEAN_weighted = tilt_method_TE_MEAN_tonnes,
    TE_rank = rank(tilt_method_TE_MEAN_weighted, ties.method = "first"),
    TE_rank_norm = scales::rescale(TE_rank, to = c(0, 1))  # normalized [0, 1]
  )

##Turn into spatial code
firm_sf <- TE_tilt_long_lat %>%
  sf::st_as_sf(coords = c("Longitude", "Latitude"), crs = 4326)

# Get Germany basemap as an sf object
germany <- ne_countries(scale = "medium", country = "Germany", returnclass = "sf")

te_ger_heatmap <- ggplot(data = germany) +
  geom_sf(fill = "gray95", color = "gray70") +
  geom_sf(data = firm_sf, aes(color = TE_rank_norm), size = 1, alpha = 0.7) +
scale_color_gradientn(
  colors = c("steelblue", "lightgray", "indianred"),
  values = scales::rescale(c(0, 0.5, 1)),
  name = "Emissions Rank\n(Normalised)"
)+
  labs(
    x = "Longitude",
    y = "Latitude"
  )

te_ger_heatmap

# Save the plot
ggsave(
  filename = "02_output_data/06_analysis/Figure-4-te_ger_heatmap.pdf",
  plot = te_ger_heatmap,
  width = 10, height = 8, dpi = 300
)

```

#04 Create plot with total emissions for top 5 sectors
Generates Figure 5:  Sectoral emissions from economy wide totals (OECD) compared to product-based (tilt)
emissions estimations, in relationship to the source’s total emissions. Expressed in percent and
weighted.

```{r}
#Join all data into one dataset
oecd_emissions_by_sector <- oecd_emissions %>%
  group_by(nace_section) |> 
  summarise(
    total_oecd_emissions = sum(total_s1_s2_s3u, na.rm = TRUE),
    total_oecd_s1 = sum(S1, na.rm = TRUE), 
    total_oecd_s2 = sum(S2, na.rm = TRUE), 
    total_oecd_s3 = sum(S3U, na.rm = TRUE),
    .groups = "drop"
  ) |> 
  mutate(
    nace_section = as.character(nace_section)
  )

# join tables and clean
comparison_emissions <- oecd_emissions_by_sector |> 
  left_join(company_emissions_by_sector, by = "nace_section") |> 
  filter(total_company_emissions > 0) |>
  mutate(
    share_of_tilt_to_oecd_emissions = (total_company_emissions/total_oecd_emissions)
  )

# add totals
totals <- comparison_emissions |>
  summarise(
    total_oecd = sum(total_oecd_emissions, na.rm = TRUE),
    total_company = sum(total_company_emissions, na.rm = TRUE)
  )

colnames(comparison_emissions)

# add share columsn
comparison_emissions <- comparison_emissions |>
  mutate(
    share_oecd_emissions =
      total_oecd_emissions / totals$total_oecd,
    share_company_emissions =
      total_company_emissions / totals$total_company
  )


colnames(comparison_emissions)

comparison_emissions <- comparison_emissions %>%
  bind_rows(
    comparison_emissions %>%
      summarise(
        across(
          where(is.numeric) & !matches("^share_of_tilt_to_oecd_emissions$"),
          sum,
          na.rm = TRUE
        ),
        share_of_tilt_to_oecd_emissions = NA_real_
      ) %>%
      mutate(nace_section = "Total") %>%
      select(nace_section, everything())
  )

colnames(comparison_emissions)

comparison_emissions <- comparison_emissions |> 
  rename(total_tilt_emissions = total_company_emissions, 
         total_tilt_s1 = total_company_s1, 
         total_tilt_s2 = total_company_s2, 
         total_tilt_s3 = total_company_s3, 
         share_tilt_emissions = share_company_emissions
  )

sector_mapper <- TE_tilt %>%
  filter(!(nace_section == "multi-sector")) %>% 
  select(nace_section, nace_section_list_adjusted) %>% 
  distinct()

comparison_emissions <- comparison_emissions %>% 
  left_join(sector_mapper, by ="nace_section")

# Prep long df (as you already do)
plot_df <- comparison_emissions |> 
  filter(nace_section != "Total") |>
  mutate(nace_label = paste(nace_section, nace_section_list_adjusted, sep = " - ")) |>
  select(nace_label, share_oecd_emissions, share_tilt_emissions) |>
  pivot_longer(
    cols = c(share_oecd_emissions, share_tilt_emissions),
    names_to = "source",
    values_to = "share"
  ) |>
  mutate(
    source = recode(
      source,
      share_oecd_emissions = "OECD emissions",
      share_tilt_emissions = "tilt emissions"
    )
  ) 

# Identify top 5 sectors by tilt share
top5 <- plot_df |>
  filter(source == "tilt emissions") |>
  arrange(desc(share)) |>
  slice_head(n = 5) |>
  pull(nace_label)

# Bucket + sum shares, then order by tilt
plot_df2 <- plot_df |>
  mutate(
    nace_group = if_else(nace_label %in% top5, nace_label, "Other Sectors")
  ) |>
  group_by(source, nace_group) |>
  summarise(share = sum(share, na.rm = TRUE), .groups = "drop") |>
  group_by(nace_group) |>
  mutate(tilt_share_for_order = share[source == "tilt emissions"][1]) |>
  ungroup() |>
  arrange(tilt_share_for_order) |>
  mutate(
    nace_group = factor(
      nace_group,
      levels = c(
        "Other Sectors",
        setdiff(unique(nace_group), "Other Sectors")
      )
    )
  ) |>
  select(nace_label = nace_group, source, share)

## PLOT
## PLOT with percentage labels
oecd_tilt_plot_share_emissions_top5 <- ggplot(
  plot_df2,
  aes(x = nace_label, y = share, fill = source)
) +
  geom_col(
    width = 0.5,
    position = position_dodge(width = 0.6)
  ) +
  # Add percent labels
  geom_text(
    aes(label = scales::percent(share, accuracy = 0.1)),
    position = position_dodge(width = 0.6),
    hjust = -0.1,  # nudges labels to the right of the bar
    size = 3
  ) +
  scale_fill_manual(
    values = c("OECD emissions" = "steelblue",
               "tilt emissions" = "indianred")
  ) +
  scale_y_continuous(
    labels = percent_format(accuracy = 0.1),
    expand = expansion(mult = c(0, 0.15))  # extra space for labels
  ) +
  labs(
    x = "Sector",
    y = "Share of total emissions",
    fill = NULL
  ) +
  coord_flip() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    strip.text = element_text(face = "bold"),
    legend.position = "top"
  )

# Save plot
ggsave(
  filename = "02_output_data/06_analysis/Figure-5-oecd_tilt_plot_share_emissions_top5_labeled.pdf",
  plot = oecd_tilt_plot_share_emissions_top5,
  width = 10, height = 8, dpi = 300
)

# Show plot
print(oecd_tilt_plot_share_emissions_top5)

```


## 05. Reduction Potential vs. Total Emissions 
Create figure with shares per sector, but color coding for absolute reduction potential
Generates Figure 6: Sectoral emission reduction potential by climate action for sectors where applicable, unweighted. 

```{r}

# --- 1. Total emissions (Scope 2 for RE1/RE4, Scope 3 for AB2) ---
sector_emissions <- TE_tilt %>%
  filter(sector_clean != "Electricity, gas, steam\nand air conditioning\nsupply") |>
  filter(tilt_method_lca_source == "ecoinvent") %>%
  mutate(
    relevant_scope_emissions = case_when(
      sector_clean == "Agriculture, forestry and\nfishing" ~ tilt_method_TE_MEAN_tonnes,
      TRUE ~ tilt_method_TE_MEAN_tonnes
    )
  ) %>%
  group_by(sector_clean) %>%
  summarise(total_emissions = sum(relevant_scope_emissions, na.rm = TRUE), .groups = "drop")

# --- 1. Reduction potentials (only relevant scopes) ---
TE_tilt_ei <- TE_tilt %>%
  filter(tilt_method_lca_source == "ecoinvent") %>%
  mutate(
    w_AB2 = if_else(sector_clean == "Agriculture, forestry and\nfishing", reduction_potential_ab2_TE_MEAN_tonnes, NA_real_),
    w_RE1 = reduction_potential_re1_TE_MEAN_tonnes,
    w_RE4 = reduction_potential_re4_TE_MEAN_tonnes
  ) |> 
  filter(sector_clean != "Electricity, gas, steam\nand air conditioning\nsupply")

reduction_sector_unweighted <- TE_tilt_ei %>%
  group_by(sector_clean) %>%
  summarise(
    AB2 = sum(w_AB2, na.rm = TRUE),
    RE1 = sum(w_RE1, na.rm = TRUE),
    RE4 = sum(w_RE4, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  pivot_longer(
    cols = -sector_clean,
    names_to = "Climate_Action",
    values_to = "Reduction_tCO2"
  ) %>%
  mutate(
    Climate_Label = case_when(
      Climate_Action == "RE1" ~ "Solar Power",
      Climate_Action == "RE4" ~ "Heat Pumps",
      Climate_Action == "AB2" ~ "Efficient Farm Machinery",
      TRUE ~ Climate_Action
    )
  ) 

unique(reduction_sector_unweighted$sector_clean)

# # --- 3. Merge emissions + compute shares ---
reduction_sector_labeled <- reduction_sector_unweighted %>%
  left_join(sector_emissions, by = "sector_clean") %>%
  filter(total_emissions > 0) %>%
  filter(!sector_clean %in% c(
    "Transportation and\nstorage",
    "Wholesale and retail\ntrade; repair of motor\nvehicles and motorcycles",
    "Accommodation and food\nservice activities", 
    "Other service activities"
  )) %>%
  mutate(
    share_percent = if_else(
      Reduction_tCO2 > 0,
      Reduction_tCO2 / total_emissions * 100,
      NA_real_
    ),
    label_text = if_else(
      Reduction_tCO2 > 0,
      paste0(round(Reduction_tCO2 / 1000, 1), " kt\n(", round(share_percent, 1), "%)"),
      NA_character_
    )
  ) %>%
  group_by(sector_clean) %>%
  filter(any(!is.na(Reduction_tCO2))) %>%
  ungroup() 

# --- 4. Plot ---
reduction_potential <- ggplot(reduction_sector_labeled, 
       aes(x = Climate_Label, 
           y = reorder(sector_clean, Reduction_tCO2), 
           fill = if_else(Reduction_tCO2 == 0, NA_real_, Reduction_tCO2)
           )
       ) +
  geom_tile(color = "white") +
  geom_text(aes(label = label_text), size = 3) +
 scale_fill_gradientn(
  colors = c("indianred", "steelblue", "seagreen"),
  values = scales::rescale(c(0, 0.02, 0.19)),
  name = "Sectoral emission reduction potential in kt CO2e",
  labels = function(x) scales::label_comma()(x / 1000),
  na.value = "white"
)+
  labs(
    x = "Climate Action",
    y = "Sector"
  ) +
  theme(
    axis.text.x = element_text(angle = 30, hjust = 1),
    plot.title = element_text(face = "bold")
  )

reduction_potential

ggsave(
  "02_output_data/06_analysis/Figure-6-reduction_potential_heatmap.pdf",
  reduction_potential,
  width = 8,
  height = 6,
  dpi = 300
)


```


#03.3.2. Quadrant Bubble
For Agriculture: 
Generates Figure 7: Emission intensity and reduction potential for industry "Crop and animal production,
hunting and related service activities" as a deep-dive, unweighted.

#03.3.2. Quadrant Bubble - for Agriculture
```{r, echo=FALSE, warning=FALSE, message=FALSE}

# 1) Prep bubble data: count duplicates of (x,y)
quad_df <- tilt_a01 %>%
  count(
    tilt_method_EF_MEAN_kg,
    tilt_method_reduction_potential_EF_MEAN_kg,
    name = "n_firms"
  ) %>%
  rename(
    x_tilt_ef = tilt_method_EF_MEAN_kg,
    y_red_ef  = tilt_method_reduction_potential_EF_MEAN_kg
  )

# 2) Quadrant cutoffs (median splits)
x_cut <- median(quad_df$x_tilt_ef, na.rm = TRUE)
y_cut <- median(quad_df$y_red_ef,  na.rm = TRUE)

# 3) Add quadrant label for later use
quad_df <- quad_df %>%
  mutate(
    quadrant = case_when(
      x_tilt_ef > x_cut & y_red_ef > y_cut ~ "upper_right",
      x_tilt_ef <= x_cut & y_red_ef > y_cut ~ "upper_left",
      x_tilt_ef <= x_cut & y_red_ef <= y_cut ~ "lower_left",
      x_tilt_ef > x_cut & y_red_ef <= y_cut ~ "lower_right"
    )
  )

# 4) Count total firms in upper right
n_upper_right <- quad_df %>%
  filter(quadrant == "upper_right") %>%
  summarise(n = sum(n_firms)) %>%
  pull(n)

# 5) Plot with annotation
p_quadrants_a01 <- ggplot(quad_df, aes(x = x_tilt_ef, y = y_red_ef, size = n_firms)) +
  geom_point(alpha = 0.7, color = "indianred") +
  geom_vline(xintercept = x_cut, linetype = "dashed", color = "grey50") +
  geom_hline(yintercept = y_cut, linetype = "dashed", color = "grey50") +

  # Add annotation to upper right quadrant
  annotate(
    "text",
    x = max(quad_df$x_tilt_ef, na.rm = TRUE) * 0.95,
    y = max(quad_df$y_red_ef, na.rm = TRUE) * 0.95,
    label = paste0("Upper-right quadrant:\n", n_upper_right, " firms"),
    hjust = 1,
    vjust = 1,
    fontface = "bold"
  ) +

  scale_x_continuous(labels = label_comma(big.mark=".", decimal.mark=",")) +
  scale_y_continuous(labels = label_comma(big.mark=".", decimal.mark=",")) +
  scale_size_area(
    max_size = 14,
    breaks = c(1, 10, 20),
    limits = c(1, 20)
  ) +
  labs(
    x = "Emission intensity (kg CO2e per EUR revenue)",
    y = "Reduction potential intensity (kg CO2e per EUR revenue)",
    size = "Number of firms"
  ) +
  theme(legend.position = "top")

p_quadrants_a01

# Save plot
ggsave(
  filename = "02_output_data/06_analysis/Figure-7-quadrants_a01.pdf",
  plot = p_quadrants_a01,
  width = 10, height = 6, dpi = 300
)


```

#Also for Chemics 
Generates Figure 13: Emission intensity and reduction potential for industry "C20 - Manufacturing of
chemicals and chemical products" as a deep-dive, unweighted. 
--> In the Appendix Section 9.4. 

```{r, echo=FALSE, warning=FALSE, message=FALSE}


# --- 1) Prep bubble data: count duplicates of (x, y) ---
quad_df <- tilt_c20 %>%
  count(
    tilt_method_EF_MEAN_kg,
    tilt_method_reduction_potential_EF_MEAN_kg,
    name = "n_firms"
  ) %>%
  rename(
    x_tilt_ef = tilt_method_EF_MEAN_kg,
    y_red_ef  = tilt_method_reduction_potential_EF_MEAN_kg
  )

# --- 2) Quadrant cutoffs (median splits) ---
x_cut <- median(quad_df$x_tilt_ef, na.rm = TRUE)
y_cut <- median(quad_df$y_red_ef,  na.rm = TRUE)

# --- 3) Count number of firms in upper-right quadrant ---
n_upper_right <- quad_df %>%
  filter(x_tilt_ef > x_cut, y_red_ef > y_cut) %>%
  summarise(total = sum(n_firms)) %>%
  pull(total)

# --- 4) Plot ---
p_quadrants_c20 <- ggplot(quad_df, aes(x = x_tilt_ef, y = y_red_ef, size = n_firms)) +
  geom_point(alpha = 0.7, color = "indianred") +
  geom_vline(xintercept = x_cut, linetype = "dashed", color = "grey50") +
  geom_hline(yintercept = y_cut, linetype = "dashed", color = "grey50") +

  # Quadrant annotation (moved up like A01)
  annotate(
    "text",
    x = max(quad_df$x_tilt_ef, na.rm = TRUE) * 0.9,
    y = max(quad_df$y_red_ef,  na.rm = TRUE) * 0.95,
    label = paste0("Upper-right quadrant:", n_upper_right, " firms"),
    hjust = 1.5,
    vjust = 1,
    size = 4,
    fontface = "bold",
    color = "black"
  ) +

scale_x_continuous(
  labels = label_comma(big.mark = ".", decimal.mark = ","),
  breaks = seq(0, 6, by = 1)
) +
coord_cartesian(xlim = c(0, 6)) +
  scale_y_continuous(labels = label_comma(big.mark = ".", decimal.mark = ",")) +
  scale_size_area(
    max_size = 14,
    breaks = c(1, 10, 20),
    limits = c(1, 20)
  ) +
  labs(
     x = "Emission intensity (kg CO2e per EUR revenue)",
    y = "Reduction potential intensity (kg CO2e per EUR revenue)", 
    size = "Number of firms"
  ) +
  theme(
    legend.position = "top"
  )

# Show plot
p_quadrants_c20

# --- 5) Save the plot ---
ggsave(
  filename = "02_output_data/06_analysis/Figure-13-quadrant_c20.pdf",
  plot = p_quadrants_c20,
  width = 10,
  height = 6,
  dpi = 300
)


```
